---
title: "Activity monitoring data"
output:
  pdf_document: default
  html_document: default
date: "2026-03-26"
---



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(lubridate)
```

1. Load and Process Data
```{r setup}
# Read dataset
data <- read.csv("C:\Users\HP Z4 G4\Documents\JHU_DS_R\activity.csv")

# Convert date column to Date type
data$date <- as.Date(data$date)

# Check structure
str(data)
summary(data)
```
2. Total Steps Per Day

steps_per_day <- data %>%
  group_by(date) %>%
  summarise(total_steps = sum(steps, na.rm = TRUE))


Histogram

ggplot(steps_per_day, aes(x = total_steps)) +
  geom_histogram(binwidth = 1000, fill = "steelblue") +
  labs(title = "Histogram of Total Steps per Day",
       x = "Total Steps",
       y = "Frequency")


Mean and Median

mean_steps <- mean(steps_per_day$total_steps, na.rm = TRUE)
median_steps <- median(steps_per_day$total_steps, na.rm = TRUE)

mean_steps
median_steps


3. Average Daily Activity Pattern

interval_avg <- data %>%
  group_by(interval) %>%
  summarise(avg_steps = mean(steps, na.rm = TRUE))

Time Series Plot

ggplot(interval_avg, aes(x = interval, y = avg_steps)) +
  geom_line(color = "blue") +
  labs(title = "Average Steps per 5-Minute Interval",
       x = "Interval",
       y = "Average Steps")


Interval with Maximum Steps

max_interval <- interval_avg %>%
  filter(avg_steps == max(avg_steps))

max_interval

4. Imputing Missing Values
Strategy

We replace missing steps values with the mean for that 5-minute interval.

# Create a lookup table for interval means
interval_means <- data %>%
  group_by(interval) %>%
  summarise(mean_steps = mean(steps, na.rm = TRUE))

# Merge and impute
data_imputed <- data %>%
  left_join(interval_means, by = "interval") %>%
  mutate(steps = ifelse(is.na(steps), mean_steps, steps)) %>%
  select(-mean_steps)


5. Histogram After Imputation

steps_per_day_imputed <- data_imputed %>%
  group_by(date) %>%
  summarise(total_steps = sum(steps))

ggplot(steps_per_day_imputed, aes(x = total_steps)) +
  geom_histogram(binwidth = 1000, fill = "darkgreen") +
  labs(title = "Histogram of Total Steps per Day (Imputed)",
       x = "Total Steps",
       y = "Frequency")

6. Weekday vs Weekend Analysis

data_imputed$day_type <- ifelse(weekdays(data_imputed$date) %in% c("Saturday", "Sunday"),
                               "Weekend", "Weekday")

interval_daytype <- data_imputed %>%
  group_by(interval, day_type) %>%
  summarise(avg_steps = mean(steps), .groups = "drop")

Panel Plot

ggplot(interval_daytype, aes(x = interval, y = avg_steps)) +
  geom_line(color = "purple") +
  facet_wrap(~day_type, ncol = 1) +
  labs(title = "Average Steps per Interval: Weekday vs Weekend",
       x = "Interval",
       y = "Average Steps")








